hatchR: A toolset to predict hatch and emergence phenology in wild fishes

### 2025-02-26 inst/manuscript/paper_pdf/paper.Rmd
added section to paper about why we don't add parameterized models to model_table

^2^US Fish and Wildlife Service, Idaho Fish and Wildlife Conservation Office, Orofino, ID, USA

@@

We demonstrate the utility of the summary and check functions `summarize_temp()`, `plot_check_temp()`, and `check_continuous()` using a simulated year-long dataset with temperature readings every thirty minutes.

```{r, echo = TRUE}
# create date object for a year with 30 min reading intervals
dates <- seq(from = ymd_hms("2000-07-01 00:00:00"),
             to = ymd_hms("2001-06-30 23:59:59"), length.out = 17568)

@@

dim(year_sim)

@@

Users can either use Salmonid models from `model_table` included in the package and the Shiny app. As discussed, these models are included because parameterizations already existed in the literature. However, these parameterizations are limited to Pacific Salmon and Bull Trout. To widen the application of the effective value approach, we include a `fit_model()` function, which is species agnostic (as long as development generally follows the power law).

>>>>

^2^Idaho Fish and Wildlife Conservation Office, U.S. Fish and Wildlife Service, Orofino, ID, USA

@@

We demonstrate the utility of the summary and check functions `summarize_temp()`, `plot_check_temp()`, and `check_continuous()` using a simulated year-long dataset (`year_sim`, first 6 rows shown) with temperature readings every thirty minutes.

```{r, echo = FALSE}
# create date object for a year with 30 min reading intervals
dates <- seq(from = ymd_hms("2000-07-01 00:00:00"),
             to = ymd_hms("2001-06-30 23:59:59"), length.out = 17568)

@@

head(year_sim)

@@

Users can either use Salmonid models from `model_table` included in the package and the Shiny app. As discussed, these models are included because parameterizations already existed in the literature. However, these parameterizations are limited to Pacific Salmon and Bull Trout (though see @quinn2018 pg. 183 for numerous other Salmonids) . To widen the application of the effective value approach, we include a `fit_model()` function, which is species agnostic (as long as development generally follows the power law). Additionally, we limit available models in `model_table` because they are well vetted models, with experimental ranges spanning 2-17 °C. Custom models we demonstrate later with `fit_model()` have much narrower parameterization ranges and, as such, we choose to keep those models out of `model_table` so as to require users to think hard about whether the custom models they are parameterizing fit the data over which they are predicting.
